---
title: "Comparison of different definitions for haemorrhagic shock"
output: html_document
---

```{r}
library(tidyverse)
library(stringr)
library(FactoMineR)

data_folder = '../../Data/'
newdata = F
if(newdata){
  traumdata <- read.csv(file=paste(data_folder,'data_trauma2_oldnames.csv', sep=''),
                        header = TRUE,na.strings = c("","NR","IMP","NA","NF"),encoding = "UTF-8")
}else{
 traumdata <- read.csv(file=paste(data_folder,'data_trauma.csv', sep=''),
                        header = TRUE,na.strings = c("","NR","IMP","NA","NF"),encoding = "UTF-8") 
 traumdata = traumdata %>% filter(!is.na(Sexe))
 traumdata$BMI[3302]=traumdata$Poids[3302]/(traumdata$Taille[3302]^2)
  traumdata=traumdata[-which(traumdata$ACR.1==1 | is.na(traumdata$ACR.1)) ,]
  traumdata=traumdata[-which(traumdata$Mecanisme=="Arme blanche" | traumdata$Mecanisme=="Arme à feu"),]
  traumdata$Origine[is.na(traumdata$Origine)] = 'Secondaire'
  traumdata = traumdata[traumdata$Origine=='Primaire',]
  traumdata= traumdata[-which(traumdata$FC.max==0),]
}

##traumdata.keep.IMP <- read.csv(file=paste(data_folder,'data_trauma.csv', sep=''),
                    #  header = TRUE,na.strings = c("","NR","NA","NF"),encoding = "UTF-8")

# If the pressure cannot be taken it means that it is low!
##traumdata$PAS[traumdata.keep.IMP$PAS=='IMP'] = 30
##traumdata$PAS.min[traumdata.keep.IMP$PAS.min=='IMP'] = 30
##traumdata$PAS.SMUR[traumdata.keep.IMP$PAS.SMUR=='IMP'] = 30
##rm(traumdata.keep.IMP)

# Recode death codes
traumdata$Cause.du.DC<-as.character(traumdata$Cause.du.DC)
traumdata[which(traumdata[ ,"Cause.du.DC"]=="Choc hémorragique"|
                  traumdata[ ,"Cause.du.DC"]=="Choc hemorragique"|
                  traumdata[ ,"Cause.du.DC"]=="Exsanguination" ), "Cause.du.DC"]<-"Exsanguination"
```

## Compute the criteria

### Def 1 : PROPPR study

```{r}
# ABC score computation
type.mecanisme.perforant = c('Arme à feu', 'Arme blanche', "Stab wound", "Gunshot wound")
mecanisme.perforant = traumdata$Mecanisme %in% type.mecanisme.perforant

lowest.PAS = pmin(traumdata$PAS.SMUR, traumdata$PAS.min, traumdata$PAS, na.rm=T) # Are values of 0 actual low pressures or no record??
low.PAS = !is.na(lowest.PAS) & (lowest.PAS <= 90) # NA values as considered as normal (same afterwards)
vasopressor = !is.na(traumdata$Dose.NAD.depart) & (traumdata$Dose.NAD.depart > 0)

highest.HR = pmax(traumdata$FC.SMUR, traumdata$FC.max, traumdata$FC, na.rm = T)
high.HR = !is.na(highest.HR) & (highest.HR >= 120)

positive.FAST = !is.na(traumdata$FAST.hemoperitoine) & (traumdata$FAST.hemoperitoine == 1)

ABC = mecanisme.perforant + (low.PAS | vasopressor) + high.HR + positive.FAST
high.ABC = ABC >= 2

# Transfusion first hour
transfusion.firstH = !is.na(traumdata$Nombre.CGR.dechoc) & (traumdata$Nombre.CGR.dechoc >= 1)

crit1 = high.ABC & transfusion.firstH
```


### Def 2 : PROMMTT study [3]:

Cannot be done with current data


### Def 3 : NAD study
```{r}
crit3 = (!is.na(traumdata$Choc.hemorragique)) & (traumdata$Choc.hemorragique == 1)
```

### Def 4 : Red flag study
```{r}
high.lactates = !is.na(traumdata$Lactates) & (traumdata$Lactates >= 5)

surgery.performed = !is.na(traumdata$Bloc.direct) & (traumdata$Bloc.direct == 1)

hemo.shock.death = !is.na(traumdata$Cause.du.DC) & (traumdata$Cause.du.DC == "Exsanguination")

crit4 = transfusion.firstH | crit3 | high.lactates | surgery.performed | hemo.shock.death
```

### Def 5 : Resuscitation Outcome Consortium criteria
```{r}
high.HR2 = !is.na(highest.HR) & (highest.HR >= 108)
HR.PAS.abnormal = high.HR2 & low.PAS
very.low.PAS = !is.na(lowest.PAS) & (lowest.PAS <= 70)
physiological.crit = HR.PAS.abnormal | very.low.PAS


unstable.pelvis = (!is.na(traumdata$Bilan.lesion.AIS)) & str_detect(traumdata$Bilan.lesion.AIS,
                             "85610.3|856151.3|856161.3|856171.4|856162.4|856163.4|856164.5|856172.4|856173.5|856174.5")
# Count positive FAST regions
positive.FAST.hemothorax = (!is.na(traumdata$FAST.hemothorax) & traumdata$FAST.hemothorax==1)
positive.FAST.epanchement = (!is.na(traumdata$FAST.epanchement.pericardique) & traumdata$FAST.epanchement.pericardique==1)
positive.FAST.hemoperitoine = (!is.na(traumdata$FAST.hemoperitoine) & traumdata$FAST.hemoperitoine==1)
positive.FAST.pneumothorax = (!is.na(traumdata$FAST.pneumothorax) & traumdata$FAST.pneumothorax==1)
multiple.positive.FAST = (positive.FAST.hemothorax + positive.FAST.epanchement +
                          positive.FAST.hemoperitoine + positive.FAST.pneumothorax) >= 2

pragmatic.crit = mecanisme.perforant | unstable.pelvis | multiple.positive.FAST

crit5 = physiological.crit & pragmatic.crit
```

```{r}
criteria = data.frame(c1=crit1, c3=crit3, c4=crit4, c5=crit5)
if(newdata){
  HS.procedure = !is.na(traumdata$procedureCH)&(traumdata$procedureCH==1)
}else{
  HS.procedure = !is.na(traumdata$Activation.procedure.CH)&(traumdata$Activation.procedure.CH==1)
}
criteria$HS.procedure = HS.procedure

traumdata = cbind(traumdata, criteria)
traumdata$Hemo.death = hemo.shock.death
```
```{r}
print('Correlation between the criteria')
cor(criteria)

causation = matrix(NA, nrow=ncol(criteria), ncol=ncol(criteria))
for(i in 1:ncol(criteria)){
  for(j in 1:ncol(criteria)){
    causation[i,j] = sum(criteria[,i]&criteria[,j])/sum(criteria[,i])
  }
}
colnames(causation) = colnames(criteria)
rownames(causation) = colnames(criteria)
print('Correlation structure (row i, column j is the proportion of times that criterion j is also true when criterion i is true)')
causation
```

```{r}
discordance = traumdata$Discordance
pred.doctor = as.character(discordance)
pred.doctor[discordance %in% c('PDSC','PGA')] = F
pred.doctor[discordance %in% c('PDAC', 'MGA')] = T
pred.doctor[is.na(discordance)] = F

criteria$doctor = pred.doctor

criteria[criteria==T] = 0.9
criteria[criteria==F] = 0.05
```


## Evaluation of the criteria

### Preliminary comparison
```{r}
# Show the relationships between the criteria
as.data.frame(table(criteria, useNA = 'ifany'))
cat('\n')
# Show the number of positives for each criterion
print(paste('Total number of positive matches (out of', nrow(traumdata), 'observations)'))
colSums(criteria)
cat('\n')
# Compute the correlation of each criteria with death by harmorrhagic shock
print('Correlation with the activation of the HS procedure')

sapply(criteria, function(x) cor(x, HS.procedure))
print('Correlation with exsanguination death')
sapply(criteria, function(x) cor(x,hemo.shock.death))
print('Correlation with overall mortality')
sapply(criteria, function(x) cor(x,!is.na(traumdata$Cause.du.DC)))
cat('\n')

print('Sensitivity (procedure)')
sapply(criteria, function(x) sum(x[HS.procedure])/sum(HS.procedure))
print('Specificity (procedure)')
sapply(criteria, function(x)sum(!x[!HS.procedure])/sum(!HS.procedure))
cat('\n')

print('Sensitivity (hemo shock death)')
sapply(criteria, function(x) sum(x[hemo.shock.death])/sum(hemo.shock.death))
print('Specificity (hemo shock death)')
sapply(criteria, function(x)sum(!x[!hemo.shock.death])/sum(!hemo.shock.death))
cat('\n')

print('Sensitivity (All deaths)')
sapply(criteria, function(x) sum(x[!is.na(traumdata$Cause.du.DC)])/sum(!is.na(traumdata$Cause.du.DC)))
print('Specificity (All deaths)')
sapply(criteria, function(x)sum(!x[is.na(traumdata$Cause.du.DC)])/sum(is.na(traumdata$Cause.du.DC)))
```
### Occurence over time
```{r}
getMonth = function(datetime){
  if(!newdata){
    t =  strptime(datetime, format = '%d/%m/%y')
  }
  else{
    t =  strptime(datetime, format = '%m/%d/%Y')
  }
  return(t$mon + 12*t$year)
}

traumdata$arrivalMonth = sapply(traumdata$Date.entree, getMonth)
traumdata %>%
  dplyr::select(arrivalMonth,c1,c3,c4,c5, HS.procedure, Hemo.death) %>%
  group_by(arrivalMonth) %>%
  summarise_all(funs(mean)) %>%
  gather('Criterion', 'Number_positives', -c(arrivalMonth)) %>%
  ggplot() + aes(x=arrivalMonth/12+1900, y=Number_positives, color=Criterion) + geom_line() +
  xlab('Time (one point per month)') + ylab('Proportion of patients matching criterion this month') + facet_grid(Criterion~.)
```

### Impact on feature selection

We ran feature selection by forward selection with the SAEM logistic regression (cf Wei's code) using each of the criteria as a target. This takes a long time to run and is not included in this document but the results are included below.

```{r}
load(paste(data_folder,'criteria.comparison.results.RData', sep=''))
bics = criteria.comparison.results$bics
aics = criteria.comparison.results$aics
best.subsets = criteria.comparison.results$best.subsets
selection.ordered = criteria.comparison.results$best.subset.ordered
print('Evolution of the BIC with the number of parameters, for each criterion')
criteria.comparison.results$bics.plot
```
```{r}
print('Subset of variable selected for the prediction of each criterion:')
best.subsets
```

```{r}
print('Order of selection of the variables for each criterion (first 8 out of a total of 14)')
selection.ordered
```

Choses à discuter (annexe):
Etablir un GUID pour les patients? De manière à facilement refaire les corrections quand la base change
De manière générale, centraliser les correctifs
