%\todo{Add graphics to explain cv}
Let us now go back to the issue of imputing missing data when there is a separation between the Practitioner and the Analyst, that is, when it is necessary to impute new incoming data without having access to the historical data.

As described Chapter \ref{imputation}, a number of methods exist to impute missing data. However, they have been designed with parameter estimation in mind: in that case, there is just one dataset that needs to be imputed. The issue, as we show in this chapter, is that current implementations of these methods may not be suitable when we need to use the same model to impute two separate datasets.

In Section \ref{ERM} we describe the process of cross-validation used to evaluate the performance that a model will have on new data. Section \ref{ERM.imp} shows how this clashes with standard implementation of imputation methods, especially when trying to compare different methods. In section \ref{ERM.solutions}, we investigate possible ways to address this issue and compare them empirically.

%On the other hand, there is almost no research on these methods when applied to prediction. Even recent manuals for machine learning \cite{ML_missdata} generally make only a quick mention of missing data, and in practice it is extremely rare for anything else than imputation by the mean to be used. \emph{Scikit-learn} \cite{scikit-learn}, by far the most-used machine learning package, only proposes imputation by the mean as of now. 

%A few research papers \cite{prediction_imputation1} \cite{prediction_imputation2} try to assess the performance of more recent imputation methods when used in a predictive context. However, they do not propose any framework or theory on this endeavour. They take it for granted that they can impute the whole dataset before performing the subsequent analysis. However, when performing prediction there is one significant difference with statistical inference, which we describe further in this chapter: the data is split into one dataset to learn the model, and another one to validate its performance. This raises many questions that we discuss here and in Chapter \ref{linreg}. 

%After laying out the general framework of Empirical Risk Minimization \cite{ERM}, which is the general paradigm used for prediction, we adapt it to fit the context of missing data. We notice that current implementations of modern imputation methods are incompatible with this framework, and try to devise solutions for this issue.
	\section{Empirical risk minimization and cross-validation}
	\label{ERM}
Let us start by ignoring the issue of missing data and assume that the data is complete. That is, we place ourselves in the same situation as described in \ref{framework} but there is no need for an Imputer, and the Analyst directly receives the data $X$.
	
As stated in Chapter \ref{data}, our end goal is to make good predictions in the real world by learning on historical data. Of course, by definition we do not have access to any future data right now, but we still need to choose a prediction and imputation model, and estimate how it will perform when we use it on the field. 

To learn and evaluate the model, we go through two steps: Empirical risk minimization (ERM) and cross-validation (CV). We describe them below
		\subsection{Empirical risk minimization}
Let us denote the $X$ a $n \times p$ matrix of covariates and $y$ a response vector of size $n$, where our goal is to predict the response $y_{new}$ from some future data $X_{new}$, assuming that $(X_{new}, y_{new})$ follow the same distribution as $X,y$. 

We choose a class of functions $f(\cdot, \beta), \beta \in B$. We want to choose a parameter $\hat{\beta}$ which we can use to predict $\hat{y}_{new} = f(X_{new}, \hat{\beta})$. The quality of a prediction is evaluated by some loss $L(y_{new},\hat{y}_{new})$. Since we do not have access to $X_{new}$, our goal is to minimize the risk: $R(\beta) = \mathbb{E}_{X_{new},y_{new}}(L(y_{new}, f(X_{new}, \beta)))$.

However, we do not know the true distribution of those values. This is why we use ERM \cite{ERM}: we define the empirical risk
$$ R_{\text{emp}}(\beta) = \frac{1}{n} \sum\limits_{i=1}^n L(y_i, f(X_i, \beta))$$
that is, the average value of the loss when predicting the known $y$ from $X$ with $\beta$.

We then select $\hat{\beta} = \argmin_{\beta} R_{\text{emp}}(\beta)$ as our ERM estimator for $\beta$. However, this is not enough. Once we have chosen $\hat{\beta}$, we need to have an estimate of how well this estimate will perform on new data.

 This is important because this is what we will take into account if we need to compare multiple choices for $f$. But the empirical risk gives us no measure of how well our model generalizes, only of how closely it can fit known data. In particular, if the class $f$ is very broad, one may find a $\beta$ that exactly interpolates the values of $y$ but does not generalize at all (an issue known as overfitting \cite{hawkins2004overfitting}).

To address the issue of model selection, we need to use CV as described below.

		\subsection{Cross-validation}
CV, consists in dividing the available data in two datasets: first we choose $n_A < n$ entries in the dataset that will be used in ERM to learn $\hat{\beta}$: this is the training dataset $X_A$ and response $y_A$. We denote $I_A = (i_1, \ldots, i_{n_A})$ the set of indices chosen for the training data and $I_V$ its complement.

The rest of the observations are noted $X_V$ and $y_V$ and called the validation dataset. They are used as a substitute for $X_{new},y_{new}$

Once this is done, the Analyst performs ERM as before, using only the training data. The obtained parameter $\hat{\beta}$ can then be evaluated with the validation error:

$$ R_{V}(\hat{\beta}) = \frac{1}{n_V} \sum\limits_{i \in I_V} L(y_i, f(X_i, \hat{\beta}))$$

It is this value that we can compare to choose the model class $f$. Once the Analyst has decided on a choice of $f$ and $\hat{\beta}$ using ERM and CV, she can send $f(\cdot, \hat{\beta})$ over to the Practitioner so that she can proceed to prediction on new data using $\hat{y_{new}} = f(X_{new}, \hat{\beta})$.

	\section{ERM with missing data: the problem of current methodologies}
	\label{ERM.imp}
We now place ourselves in the same context as before, except some values are missing from $X$, both in the training and the validation data. This means that we are back to a case where there is an Imputer in addition to the Analyst.
		\subsection{Imputation seen as an ERM}
Remember that the purpose of this work is to impute the data independently of the model used afterwards for prediction. This means that we cannot perform ERM exactly as before and use any function we like to go from $X$ (which has missing data) to $\hat{y}$. The prediction is the composition of two steps.
			\paragraph{Imputation step}
First we choose an imputation model $\hat{X} = g(X, \alpha)$ where $\hat{X}$ is the completed dataset and $\alpha$ some parameter. This is similar to the previously described ERM, except we do not know the true data (while we had $y$ to compare to $\hat{y}$, we do not know the true full dataset $\tilde{X}$). Thus, we choose $\hat{\alpha}$ to minimize some unsupervised loss
$$L'(g(X, \alpha), \alpha)$$
The loss $L'$ would usually be the negative log-likelihood of a distribution model for $X$. For example, we can choose to perform imputation by maximum likelihood using a normal approximation. In that case, we would have $alpha =(\mu, \Sigma)$ and \cite{ref_amelia}
$$L'(\hat{X}, \alpha) = \phi(\hat{X};\mu, \Sigma) $$ (where $\phi$ is the p.d.f.\ of the normal distribution)
with $\hat{X} = \argmax_{A} \phi(A;\mu, \Sigma)$ (in this case, this corresponds the expected value of $X$).

			\paragraph{Prediction step}
With imputation done, we can proceed as before to choose a parameter $\hat{\beta}$ that minimizes the empirical risk when using the completed data:
	$$ R_{\text{emp}}(\beta, \hat{X}) = \frac{1}{n} \sum\limits_{i=1}^n L(y_i, f(\hat{X}_i, \beta))$$
	
Putting it all together, we can define (with a slight abuse of notation, we use $f(\hat{X}, \beta)$ to refer to function $f$ applied to each line of $\hat{X}$):
$$ h(X, (\alpha, \beta)) = f(\hat{X}, \beta) = f( g(X, \alpha), \beta) $$
the combined model that takes the observed data as input and outputs a predicted $y$. Formally, the two successive steps yield:
\begin{align*}
\hat{\alpha} &= \argmin_{\alpha} L'(g(X,\alpha), \alpha) \\
\hat{\beta} &= \argmin_{\beta} R_{\text{emp}}(\beta,g(X,\hat{\alpha})) \\
\hat{y} &= h(X, (\hat{\alpha}, \hat{\beta}))
\end{align*}

We choose to use this notation to illustrate our point that imputation is an integral part of the ERM, not a separate, preliminary process. In particular, it means that its parameters must be subjected to CV just like those of the prediction. That is, only $X_A$ and $y_A$ are used to estimate $(\hat{\alpha}, \hat{\beta})$ as shown above, while $X_V$ and $y_V$ are held out. Then, we can compute a prediction $\hat{y}_V = h(X_V, (\hat{\beta}, \hat{\alpha}))$ and we compute $L(y_V, \hat{y}_V)$ to evaluate the choice of model.

This implies that just like $\beta$, the imputation parameter $\alpha$ should be estimated only on the training data and then used to impute and predict the validation data. As we see below, this particular requirement clashes with standard implementations of imputation methods.

		\subsection{Implementation issues}
Current standard implementations of imputation methods are usually used through a single function which takes a dataset with missing values as an input and returns a completed dataset, but not simple a way to impute a new dataset with the exact same parameters. \cite{stekhoven2015missforest} \cite{josse2016missmda}\cite{MICE_founding}\cite{ref_amelia}

Although this is fine for most applications, in the case of CV this is problematic. To perform CV properly, one would estimate $(\hat{\alpha}_A, \hat{\beta}_A)$ through ERM, and then make a prediction on the validation set as $h(X_V,(\hat{\beta}_{X_A},\hat{\alpha}_{X_A}))$. But if all one has access to is a black-box function $g': X \mapsto g(X, \hat{\alpha}_X)$, where $\hat{\alpha}_X$ is the optimised parameter for the argument $X$, then one cannot choose what parameters are used to impute the input to function $g'$: a new parameter will be estimated at every call of the function. But in that case, it is impossible to use the same $\alpha$ for the training dataset and for the validation data --- or for new data.

If we are to use one of these methods to build a tool for hemorrhagic shock, we need to be able to impute on a new incoming patient, without any access to the training data or to other new patients' data, so we have no choice but to reuse previously computed parameters: a reimplementation is unavoidable.

In some cases it is straightforward to reimplement a given method to separate the parameter estimation and the imputation. However, in other cases (e.g.\ iterative methods such as PCA imputation) this is a much more involved task. In particular, one may want a way to rapidly compare many imputation methods using existing software packages, before choosing one to reimplement. 

This is why we are interested in alternatives that could be used to impute the data using only a black-box function, to use as a proxy of the performance of the properly reimplemented method. We want to see if there is a way to order the relative performance of various imputation methods using such a proxy, before choosing one to implement.

	\section{Possible solutions}
	\label{ERM.solutions}
	
If we were to follow exactly the principles of CV, we would proceed as follows: 
\begin{algorithm}[H]
	\caption{Separate imputation}
	\hspace*{\algorithmicindent} \textbf{Input:} $X$, $y$, $I_A=\{i, X_i \in X_A\}$ \\
 	\hspace*{\algorithmicindent} \textbf{Output:} $\hat{y}_V$
	\begin{algorithmic}[1]
		\State \textbf{Parameter estimation:}
		\Indstate $\hat{\alpha}_A \leftarrow \argmin_{\alpha} L'(g(X_A,\alpha)$
		\Indstate $\hat{X}_A \leftarrow g(X_A, \hat{\alpha}_A)$
		\Indstate $\hat{\beta}_A \leftarrow \argmin_{\beta} R_{\text{emp}}(\beta,\hat{X}_A)$
		\State \textbf{Prediction:}
		\Indstate $\hat{X}_V \leftarrow g(X_V, \hat{\alpha}_A)$ \Comment \emph{Uses the same $\alpha$ as for the training set}
		\Indstate $\hat{y}_V \leftarrow f(\hat{X_V}, \hat{\beta}_A)$
	\end{algorithmic}
\end{algorithm}

We call this Separate Imputation (SI). But this is not possible using a black-box function because we need to recover $\hat{\alpha}_A$ and reuse it with $X_V$.

		\subsection{Alternatives using current implementations}
			\subsubsection{Methods}

\paragraph{Grouped imputation (GI)} 

A way to perform the imputation with just one function is to impute all the data at once before performing the CV split:
\begin{algorithm}[H]
	\caption{Grouped imputation}
	\hspace*{\algorithmicindent} \textbf{Input:} $X$, $y$, $I_A=\{i, X_i \in X_A\}$ \\
 	\hspace*{\algorithmicindent} \textbf{Output:} $\hat{y}_V$
	\begin{algorithmic}[1]
		\State \textbf{Imputation:}
		\Indstate $\hat{X} \leftarrow g'(X) = g(X, \hat{\alpha}_X)$ 
		\Indstate $(\hat{X}_A, \hat{X}_V) \leftarrow \hat{X}$ \Comment \emph{Split the data after imputation}
		\State \textbf{Estimation of the prediction parameter}
		\Indstate $\hat{\beta}_A \leftarrow \argmin_{\beta} R_{\text{emp}}(\beta,\hat{X}_A)$ \Comment \emph{Note that $\hat{X}_A = g(X_A, \hat{\alpha}_X)$}
		\State \textbf{Prediction:}
		\Indstate $\hat{y}_V \leftarrow f(\hat{X_V}, \hat{\beta}_A) = f(g(X_V, \hat{\alpha}_X)$
	\end{algorithmic}
\end{algorithm}

Here, both datasets are imputed with the same parameter $\alpha_X$ but this means that the validation data is used to choose that parameter which then serves to impute the training data. In that sense, the CV does not faithfully reproduce the actual real-world application where the new data would be unavailable at the time of training --- and the training data unavailable at the time of prediction. As a consequence the validation error may not be representative of real-world performance and could modify the performance order when comparing multiple methods.

\paragraph{Independent imputation (II)}
Another approach is to divide the data first, then impute each dataset independently:
\begin{algorithm}[H]
	\caption{Independent imputation}
	\hspace*{\algorithmicindent} \textbf{Input:} $X$, $y$, $I_A=\{i, X_i \in X_A\}$ \\
 	\hspace*{\algorithmicindent} \textbf{Output:} $\hat{y}_V$
	\begin{algorithmic}[1]
		\State \textbf{Training parameter estimation:}
		\Indstate $\hat{X}_A \leftarrow g'(X_A) = g(X_A, \hat{\alpha}_A)$
		\Indstate $\hat{\beta}_A \leftarrow \argmin_{\beta} R_{\text{emp}}(\beta,\hat{X}_A)$
		\State \textbf{Imputation of $X_V$:}
		\Indstate $\hat{X}_V \leftarrow g'(X_V) = g(X_V, \hat{\alpha}_V)$ \Comment \emph{Imputation made independently of that of $X_A$}
		\State \textbf{Prediction:}
		\Indstate $\hat{y}_V \leftarrow f(\hat{X_V}, \hat{\beta}_A)$
	\end{algorithmic}
\end{algorithm}

Here, we respect the rules of CV (the validation data is not used at all during training), but we are using parameter $\hat{\alpha}_V$ to impute $X_V$, while we learned $\hat{\beta}_A$ on $\hat{X}_A$ which was imputed with $\hat{\alpha}_A$. That is, we are optimising $h(\cdot,(\hat{\alpha}_A, \hat{\beta}_A))$ and predicting with $h(\cdot,(\hat{\alpha}_V, \hat{\beta}_A))$. As for GI, this could have an impact on the validation error.

$\hat{\alpha}_V$, $\hat{\alpha}_A$ and $\hat{\alpha}_X$ are asymptotically the same for large $n$ --- $X_A$, $X_V$ and $X$ have the same distribution since the lines of $X$ are i.i.d\ ---, so all three methods (Separate, Grouped and Independent) should be identical for large $n$. But for smaller $n$, harmful effects may be present --- overoptimistic validation error due to 'cheating' for GI, high error due to the difference in parameters for II.

%\paragraph{Line by line imputation}
%The last option we mention here is to first impute the training data on its own. Then for every line of the validation data we impute it by stacking it with the imputed training data and imputing the whole dataset. Since it is just one line, we can safely assume that the imputation parameters will be those of the training data. The main issue with this method is that is the validation data is rather large, this will be computationally infeasible.

			\subsubsection{Need for a new implementation}
We want to understand whether the alternatives proposed here are good enough to be used if SI is not available. To do that, we need at least one implementation where SI is available. That way we will be able to compare its performance with GI and II.

Moreover, in addition to this theoretical pursuit, we need this because of what we are trying to achieve with Traumabase: the end goal is to make a recommendation system that can produce a prediction for \emph{a single new patient} arriving to the hospital, without needing to have access to the whole Traumabase data. Without access to the initial training data, this means that only a fully parametric approach can be taken in this particular case (II is impossible on just one line of data, and GI requires access to the full data).

Below, we design a very simple imputation method for those purposes.

		\subsection{Multivariate normal conditional expectation}
The principle of this imputation is inspired from R package \emph{Amelia} \cite{ref_amelia}, and a large part of the code is from the \emph{norm} package \cite{pkg_norm}. The idea is to model both $X_A$ and $X_V$ as normally distributed $\mathcal{N}(\mu, \Sigma)$ with unknown parameters. Although this is often a rough approximation of the true distribution, it has been shown to  perform well on a range of datasets \cite{MICE_founding}\cite{schafer1998multiple}. In the Traumabase data, the variables are unimodal and have almost symmetric distributions so we can expect a normal approximation to be resonable.

This hypothesis allows us to impute the missing data conditionally on the observed data.

\paragraph{Parameter estimation}
It is possible to approximate maximum-likelihood estimators for $\mu$ and $\Sigma$ with an iterative procedure, using the EM (expectation-maximisation) algorithm \cite{EM}. The algorithm is detailed in \cite[Chapter 5.3]{em_normal_fit}
%\begin{algorithm}[H]
%	\caption{Normal parameter estimation with EM}
%	\hspace*{\algorithmicindent} \textbf{Input:} $X$ \\
% 	\hspace*{\algorithmicindent} \textbf{Output:} $\hat{\mu}, \hat{\Sigma}$
%	\begin{algorithmic}[1]
%		\State $X^{(0)} \leftarrow X$ where missing values are replaced using the observed mean of $X$ (mean imputation)
%		\State $t \leftarrow 0$
%		\While {not converged}
%			\State $(\mu^{(t)}, \Sigma^{(t)}) \leftarrow$ sample mean and covariance matrix of $X^{(t)}$ \Comment \emph{i.e.\ maximum likelihood %estimates on the completed data}
%			\State $X^{(t+1)} \leftarrow \mathbb{E}_{X^{\text{miss}}}(X \vert X^{\text{obs}} ; \mu^{(t)}, \Sigma^{(t)})$ \Comment \emph{Missing values are replaced by their expected value under the new parameters}
%			\State $t \leftarrow t+1$
%		\EndWhile
%		\State $\hat{\mu}, \hat{\Sigma}$ = $\mu^{(t)}, \Sigma^{(t)}$
%	\end{algorithmic}
%\end{algorithm}

 For this step, we use a slightly modified version of the code from the \emph{norm} package.

\paragraph{Imputation}
Once we have the parameters, it is very straightforward to get an imputation of the missing data. We impute using the conditional expectation of the dataset conditioned on the observed values: 
$$\hat{X} = \mathbb{E}_{X^{\text{miss}}}(X \vert X^{\text{obs}} ; \hat{\mu}, \hat{\Sigma})$$

The conditional expectations are easily derived using the Schur complement \cite{norm_schur}.

We implemented this step as well to get a complete imputation procedure divided in two functions that implement the estimation and imputation steps separately. The code is available in package SIMVN \cite{SIMVN_github}.

In all that comes next, unless specified otherwise this imputation method is the one we use.
		\subsection{Comparison on simulated data}
Now that we have an implementation that separates estimation and imputation, we use it to compare the three imputation procedures on simulated data (cf Appendix \ref{simulation}, with $\rho=0.5, p=4, \sigma=1$) and the abalone data (cf Appendix \ref{abalone}) with various sample sizes and adding 30\% MCAR missing data. The results are in Figure \ref{fig.grouped_separate}.

\input{Graphs/fig_grouped_separate}
	
We see that indeed GI seems to always have lower error than SI, while II has higher error. However, when $n$ is not too small, the relative difference between GI and SI is small. II seems to be further off than the other two, and sometimes has high error. As a results, GI appears to be a good proxy for SI performance: it can be used a a quick check to compare the performances of a range of imputation methods, in order to choose one method to reimplement for the final SI imputation.
